---
title: "Lab 2 Team File"
author: "Mitchell Kazin and Adam VanGorp"
date: "9/19/2019"
output: html_document
---
Link: https://github.com/mitchellkazin15/ds202_lab2 <br>
Owner username: mitchellkazin15 <br>
Repository: ds202_lab2 <br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Adam VanGorp
Mitch Kazin

https://github.com/mitchellkazin15/ds202_lab2.git


```{r}
  choco <- read.csv("https://raw.githubusercontent.com/lbchu/ds202_labs/master/choco.csv")
  head(choco)
```

## Problem 1

```{r,eval=FALSE}
  recent_rating = choco[choco$Review.Date>2011,] # grabs data that is from 2011 and onward
  unique(recent_rating$Company.Location) # NOTICE: how many countries are listed but how many levels are there? 
  recent_rating$Company.Location = droplevels(recent_rating$Company.Location) #gets rid of unused levels in the new data
  head(recent_rating) #displays first few rows of the new dataset
  
  cross_table = table(recent_rating$Review.Date,recent_rating$Company.Location) #organizes new data by location and review date
  cross_table #displays the table created
  which(cross_table[1,]!=0 & cross_table[2,]!=0 & cross_table[3,]!=0 & cross_table[4,]!=0 & cross_table[5,]!=0) # only shows rows that have value in every column
  
  france_recent = subset(recent_rating,Company.Location =='France') #grabs only France data
  uk_recent = subset(recent_rating,Company.Location =='U.K.') #grabs only UK data
  germany_recent = subset(recent_rating,Company.Location=='Germany') #grabs only Germany data
  
  par(mfrow=c(2,2)) # sets number of sublplots that will be displayed
  
  color1 = ifelse(france_recent$Rating >median(france_recent$Rating), 'blue', 'green') #colors higher ratings blue and lower ratings green
  
  plot(france_recent$Cocoa.Pct ~ france_recent$Review.Date,ylim=c(50,90),col=color1,pch=18,cex=1.5,xlab="Review Date", ylab= "Cocoa Percentage", main= "France") #plots the France data
  legend('bottomright',legend=c('High Rating', 'Low Rating'),col=c('blue','green'),pch=18) #adds legends to explain the color
  
  color2 = ifelse(uk_recent$Rating >median(uk_recent$Rating), 'purple', 'red') #colors higher rtings purple and lower ratings red
  plot(uk_recent$Cocoa.Pct ~ uk_recent$Review.Date,ylim=c(50,90),col=color2,pch=18,cex=1.5, xlab="Review Date", ylab= "Cocoa Percentage", main= "UK") #plots UK data
  legend('bottomright',legend=c('High Rating', 'Low Rating'),col=c('purple','red'),pch=18) #adds legend
  
  color3 = ifelse(germany_recent$Rating > median(germany_recent$Rating), 'orange', 'yellow') #colors high ratings orange and lower yellow
  plot(germany_recent$Cocoa.Pct ~ germany_recent$Review.Date,ylim=c(50,90),col=color3,pch=18,cex=1.5,xlab="Review Date", ylab= "Cocoa Percentage", main= "Germany") #rmany data g
  legend('bottomright',legend=c('High Rating', 'Low Rating'),col=c('orange','yellow'),pch=18) #adds legend
```
## Problem 2

The colors used for High and Low Rating should be the same for each visual.

## Problem 3

```{r}
  rating_counts = table(choco$Review.Date)
  barplot(rating_counts,xlab="Review Date", ylab= "Ratings", main= "Ratings by Year")
```

There has been a steady increase in ratings from 2006 to 2015 but then there's a steep decline.

## Problem 4

``` {r}
rating_counts = table(choco$Company.Location)
rating_counts = subset(rating_counts, rating_counts > 30)
sorted_rating_counts = sort(rating_counts, decreasing = TRUE)
sorted_rating_counts
USA = subset(choco$Rating, choco$Company.Location == 'U.S.A.')
France = subset(choco$Rating, choco$Company.Location == 'France')
Canada = subset(choco$Rating, choco$Company.Location == 'Canada')
USA_avg_rating = mean(USA)
France_avg_rating = mean(France)
Canada_avg_rating = mean(Canada)
USA_avg_rating
France_avg_rating
Canada_avg_rating                      
```
The average rating for the three countries with the most reviews are very similiar with the nice Canadians giving the highest average rating. 

## Problem 5

What companies that have over 20 reviews have the highest average ratings?

## Problem 6

```{r}
  rating_counts = table(choco$Company)
  rating_counts = subset(rating_counts, rating_counts > 20)
  rating_counts
  top_companies = subset(choco[c(1,7)],choco$Company %in% names(rating_counts))
  avg_rating = aggregate(top_companies$Rating,  by = list(unique.values = top_companies$Company), FUN=mean)
  barplot(avg_rating$x, names.arg = avg_rating$unique.values, xlab="Company", ylab= "Average Rating", main= "Average Rating by Top Companies", ylim = c(0,5), cex.names = .9)
```

All companies have a close average rating of around 3.5 with Soma being the highest rated and Guittard being the lowest.

